---
title: 'Obtaining an lx mx schedule from a matrix'
date: "6 Oct 2015"
output: pdf_document
classoption: a4paper
---
  

It is possible to obtain `lx` and `mx` schedules from matrix population models using the methods described by Caswell (2001).



```{r echo=FALSE}
load("~/Dropbox/COMADRE intro ms/Data/COMADRE_v.1.0.0.RData")
```

Load the data:
```{r eval=FALSE}
load("COMADRE_v.1.0.0.RData")
```

Next, query the `metadata` part of `comadre` to identify the location of the desired data using the command `which` in a conditional search:

```{r comment=">"}
id <- which(comadre$metadata$SpeciesAccepted == "Lepus_americanus")
id
```

The code to do the mathematics is available at our GitHub site.

You will first need to load the `devtools` library, and also `MASS` since the function calls functions in `MASS`:

```{r}
library(devtools)
library(MASS)
```

Now load the required functions:

```{r echo = TRUE}
source_url("https://raw.githubusercontent.com/jonesor/compadreDB/master/Functions/makeLifeTable.R")
```

Let's pick the first of the list obtained above:

```{r comment = ">"}
lt1 <- makeLifeTable(matU = comadre$mat[[id[1]]]$matU, matF = comadre$mat[[id[1]]]$matF ,startLife = 1, nSteps=10)
lt1
```


  