---
title: "Geographic distribution"
date: "`r Sys.Date()`"
output: pdf_document
classoption: a4paper
---

This example produces on a world map the viability (population growth rate $\lambda$ > 1, $\lambda$ = 1, $\lambda$ < 1) of a subset of studied populations given some selection criteria, and color-codes the location of each population according to the value of $\lambda$.

First, subset mean matrices for all Carnivora in the wild in the northern hemisphere, with no issues for survival >1, for which matrices have been split into **A** = **U** + **F** + **C**, and for which reproduction was explicitly modeled.


```{r echo=FALSE}
load("~/Dropbox/COMADRE intro ms/Data/COMADRE_v.1.0.0.RData")
```

Load the data:
```{r eval=FALSE}
load("COMADRE_v.1.0.0.RData")
```


```{r, comment=">"}
tempMetadata <- subset(comadre$metadata,
                       MatrixComposite == "Mean" &
                         Order == "Carnivora" &
                         MatrixCaptivity == "W" &
                         LatNS == "N" &
                         SurvivalIssue < 1 &
                         MatrixSplit == "Divided" &
                         MatrixFec == "Yes")
```

 The row names from the subsetted dataframe can now be used to subset the entire `comadre` database using the function `subsetDB` which is available via GitHub (using `source_url` from `devtools` to download it).

```{r}
library(devtools)
source_url("https://raw.githubusercontent.com/jonesor/compadreDB/master/Functions/subsetDB.R")
```


```{r, comment=">"}
id <- as.numeric(rownames(tempMetadata))
x<-subsetDB(comadre,id)
```

The object `x` is now a version of the `comadre` database object that contains only the matrices that match the search criteria. To calculate population growth rate for the subset matrices, we can first create an empty `data.frame` to accommodate the output:

```{r, comment=">"}
output <- data.frame(lambdas = rep(NA, length(x$mat)))
```

Now create dummy variables to convert geographic information to be plotted on the map (below):

```{r, comment=">"}
x$metadata$LAT <- NA
tempMetadata$LON <- NA

for(i in 1:nrow(x$metadata)){
  if(x$metadata$LatNS[i] == "S") {
    x$metadata$LatDeg[i] <- -x$metadata$LatDeg[i]
  }
  if(x$metadata$LonWE[i] == "W") {
    x$metadata$LonDeg[i] <- -x$metadata$LonDeg[i]
  }
  x$metadata$LAT[i] <- x$metadata$LatDeg[i] + 
    x$metadata$LatMin[i]/60 + x$metadata$LatSec[i]/3600
  x$metadata$LON[i] <- x$metadata$LonDeg[i] + 
    x$metadata$LonMin[i]/60 + 
    x$metadata$LonSec[i]/3600
}
```

Create an empty variable to accommodate output from lambda calculations:

```{r, comment=">"}
x$metadata$lambdas <- NA
```


Then, create a `for` loop to examine each matrix in turn. Here it may be advisable to use the function `tryCatch` as a wrapper to cope with the situation if/when the function in the loop fails:

```{r, comment=">"}
for (i in 1:length(x$mat)){
  tryCatch({
    x$metadata$lambdas[i] <- Re(eigen(x$mat[[i]]$matA)$value)[1]
  }, error = function(e){})
}
```

Now we can create a vector of color hex codes that can be applied according to the estimate of $\lambda$. THis is done using the `colorRampPalette` function to go from green for high values of $\lambda$, to red for low values of $\lambda$. Here `paste` is used to append a value of `90` to the hex codes to allow transparency for aesthetic reasons.

```{r, comment=">"}

rampfunc <- colorRampPalette(c("green", "red"))
colVect <- rampfunc(100)
colVect <- paste(colVect,"90", sep="")
s1 <- seq(min(x$metadata$lambdas, na.rm=TRUE),max(x$metadata$lambdas, na.rm=TRUE),
          length.out = 100)
```

First, load the `maps` package (and install it if necessary). Then plot the world map and overlay the points from our data, color coded by value of $\lambda$. In this case, the points are jittered slightly to improve visibility of nearby populations.

```{r, comment=">"}
library(maps)
map("world", col = "gray", fill = TRUE, bg = "light blue",
    xlim = c(-175, 176), ylim = c(-60, 85), border = "white")

points(jitter(x$metadata$LON, 0.6), jitter(x$metadata$LAT, 0.6),
       col = colVect[findInterval(x$metadata$lambdas, s1)],cex = 2, pch = 16)
```