---
title: "Obtaining an lx mx schedule from a matrix"
date: "`r Sys.Date()`"
output: pdf_document
classoption: a4paper
---
  

It is possible to obtain `lx` and `mx` schedules from matrix population models using methods described by Caswell (2001). There is a function `makeLifeTable` in our `Mage` package, that implements this.

Here is an example.



```{r echo=FALSE}
load("~//Dropbox/MatrixDatabases/COMADRE/v.2.0.0/COMADRE_v.2.0.0.RData")
library(Mage)
library(MASS)
```

Load the data:
```{r eval=FALSE}
load("COMADRE_v.2.0.0.RData")
```

Next, query the `metadata` part of `comadre` to identify the location of the desired data using the command `which` in a conditional search:

```{r comment=">"}
id <- which(comadre$metadata$SpeciesAccepted == "Lepus_americanus")
id
```


Let's pick the first matrix of the list obtained above:

```{r comment = ">"}
lt1 <- makeLifeTable(matU = comadre$mat[[id[1]]]$matU, matF = comadre$mat[[id[1]]]$matF,
                     startLife = 1, nSteps=10)
lt1
```


  