---
title: "phyloExample"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data

```{r echo=FALSE}
load("~/Dropbox/MatrixDatabases/COMADRE/v.2.0.0/COMADRE_v.2.0.0.RData")
library(Mage)
library(ape)
library(caper)
library(popbio)
```

We will make use of the mammalian supertree.

```{r, comment=">"}
mamtree <- read.nexus("/Users/jones/Dropbox/ESA\ 2017\ Portland\ Workshop/data/mammaltree.nex")
summary(mamtree)
```


## Subsetting the data

Firstly, subset to mean matrices, for mammals, that have a duration of 3+ and dimension of >3.

```{r, comment=">"}
x<-subsetDB(comadre, MatrixComposite == "Mean" &
                         Class == "Mammalia" &
                         StudyDuration >= 3 &
                         MatrixDimension > 3)
```

The straightforward phylogenetic regression models require one measure per species.
So I will take only the first entry for each species if there are duplicates.
In addition, I want to subset to only those species that are actually in our phylogenetic tree.

```{r, comment=">"}
temp <- which(!duplicated(x$metadata$SpeciesAccepted))
rownames(x$metadata)<-NULL
y<-subsetDB(db=x,sub = rownames(x$metadata) %in% temp & 
              SpeciesAccepted %in% mamtree$tip.label)
```
So now we have a version of COMADRE called `y` that contains only the species of interest.

Now we can loop through those species and make some calculations.
Here I am calculating life expectancy from the start of the first stage, which is the sum of the first column of the fundamental matrix.

This code looks a little complicated because it is not possible to calculate the fundamental matrix for some matrices. Therefore I have to add an `if` statement to cope with those problems.

```{r, comment=">"}
demog <- data.frame(species = y$metadata$SpeciesAccepted)
demog$lexp <- NULL
for(i in 1:nrow(y$metadata)){
  A <- y$mat[[i]]$matA
  fm <- fundamental.matrix(A)
  if(class(fm) == "list"){
  demog$lexp[i]<-sum(fm$N[,1])}else{demog$lexp[i] <- NA}
}
```

Now I have a data frame called `demog` that includes the species name and life expectancy.

Now I can import weight data for these species.

```{r, comment=">"}
mammweight<-read.csv("/Users/jones/Dropbox/ESA\ 2017\ Portland\ Workshop/data/mammalweight.csv")
```

And merge it with the existing data.

```{r, comment=">"}
demog2 <- merge(demog,mammweight,by="species")
demog2$logmass <- log(demog2$weight)
demog2$loglexp <- log(demog2$lexp)
```

I can now plot the relationship between log mass and log life expectancy.

```{r, comment=">"}
plot(demog2$logmass,demog2$loglexp)
```

For the next bit we will use functions from `caper`.

```{r, comment=">"}
library(caper)
```

We make a "comparative data object"

```{r, comment=">"}
mamdat <- comparative.data(data = demog2, phy = mamtree, names.col = "species",na.omit=FALSE)
```

Fit a PGLS model of the relationship between longevity and weight.

```{r, comment=">"}
mod1<-pgls(loglexp~logmass,mamdat,lambda="ML")
summary(mod1)
pgls.profile(mod1)
plot(pgls.profile(mod1))
```

Let's plot this regression, and compare it to an ordinary least squares regression.

```{r, comment=">"}
plot(demog2$logmass,demog2$loglexp)
abline(mod1)


modols<-lm(loglexp~logmass,mamdat$data)
abline(modols,col="red")
```


