---
title: "Plot a life cycle diagram from a matrix model"
date: '`r Sys.Date()`'
output: pdf_document
classoption: a4paper
---
  

The exercise plots a life cycle with the stages and transitions of a give matrix chosen from the `comadre` database. It will use the R function `plotLifeCycle` from the COMADRE/COMPADRE GitHub repository. This function works well with matrices of relatively low dimensionality (~< 7), and where not many transitions are depicted. The function uses the `DiagrammeR` package, so this needs to be called first:

```{r, comment=">"}
library(DiagrammeR)
```

Let's plot the lifecycle described by the `A` matrix for a species containing the word "lion" in the common name used by the author(s) in the original source publiction.

```{r echo=FALSE}
load("~/Dropbox/COMADRE intro ms/Data/COMADRE_v.1.0.0.RData")
```

First load the data:
```{r eval=FALSE}
load("COMADRE_v.1.0.0.RData")
```

To find the species with the word `lion` in their common name, we use the function `grep`.

```{r, comment=">"}
lions <- grep("lion",comadre$metadata$CommonName)
unique(comadre$metadata$CommonName[lions])
```

Unfortunately, no actual lion (*Panthera leo*) has been included in this version of COMADRE, but there are plenty of other "lions" in it. We will plot the life cycle of the red lionfish (*Pterois volitans*):

```{r, comment=">"}
matNum <- which(comadre$metadata$CommonName == "Red lionfish")
matNum
sp <- gsub("_"," ",comadre$metadata$SpeciesAccepted[matNum])
sp
```

The matrix 'A' and the stages of this study are:

```{r, comment=">"}
matA <- comadre$mat[[matNum]]$matA
matA
stages <- comadre$matrixClass[[matNum]]$MatrixClassAuthor
stages
```

To plot its lifecycle, use the function `plotLifeCycle.R` from the `Mage` package. 


```{r eval =FALSE}
plotLifeCycle(matA, title = "Red lionfish")
```

