---
title: 'Simple outputs from many matrices'
date: "`r Sys.Date()`"
output: pdf_document
classoption: a4paper
---


This example produces some basic output such as the population growth rate ($\lambda$) and damping ratio (Caswell 2001) for a subset of species and populations given some selection criteria.

```{r echo=FALSE}
load("~/Dropbox/MatrixDatabases/COMADRE/v.2.0.0/COMADRE_v.2.0.0.RData")
library(Mage)
```

First, use `subsetDB` (from the `Mage` package) to subset the database to the data of interest: only mean matrices for bony fish from studies of 3 years duration or longer, and with a  matrix dimension of 3 or greater.

```{r, comment=">"}
x<-subsetDB(comadre, MatrixComposite == "Mean" &
                         Class == "Actinopterygii" &
                         StudyDuration >= 3 &
                         MatrixDimension > 3)
```

The object `x` is now a version of the `comadre` database object that contains only the matrices that match the search criteria.

These matrices can now be analyzed by applying functions in a loop, or by using `lapply`.

For example, to calculate population growth rate and damping ratio for
the subset of matrices, first create an empty `data.frame` to accommodate the output

```{r, comment=">"}
output <- data.frame(lambdas = rep(NA, length(x$mat)),
                     damps = rep(NA, length(x$mat)))
```

and then use the functions in `popbio` package to derive demographic output (the user may need to install the package first).

```{r, comment=">"}
library(popbio)
```


```{r, comment=">"}
for (i in 1:length(x$mat)){
    output$lambdas[i] <- Re(eigen(x$mat[[i]]$matA)$value)[1]
    output$damps[i] <- damping.ratio(x$mat[[i]]$matA)
}
```

To examine the output:

```{r, comment=">"}
output
```

These parameters are presented in the same order as the `metadata` so we could also add the species name to this `data.frame`:

```{r, comment=">"}
data.frame(Species = x$metadata$SpeciesAccepted,output)
```

Plot the population growth rates and damping ratios derived from these matrices. In this plot, the vertical, dashed red line indicates 
 population growth rate = 1 (or log ($\lambda$) = 0)

```{r, comment=">"}
par(mfrow = c(1,2))
hist(log(output$lambdas),
     xlab = "Log population growth rate",
     col = "gold", main = "")
abline(v=0,col = "red", lwd = 4, lty = 3)
hist(output$damps, xlab = "Damping ratio",
     col = "brown", main = "")
```
