---
title: "Convergence to QSD"
date: "`r Sys.Date()`"
author: Owen Jones
output: pdf_document
classoption: a4paper
---


```{r echo=FALSE}
load("~/Dropbox/COMADRE intro ms/Data/COMADRE_v.1.0.0.RData")
```

This uses a function from the GitHub repository:
```{r,message=FALSE}
library(devtools)
source_url("https://raw.githubusercontent.com/jonesor/compadreDB/master/Functions/QSDConverge.R")
```

You will also need the `popbio` package.
```{r,message=FALSE}
library(popbio)
```

First, select the U matrix subpart of the matrix you are interested in.

```{r}
m1<-comadre$mat[[655]]$matU
```

Second, identify the first "active" stage where you'd normally consider life to begin. 
This is mostly important for species with "seed" stages, or other kinds of dormant stages.

```{r}
SL <- which(comadre$matrixClass[[655]]$MatrixClassOrganized == "active")[1]
```

Finally use the `qsdConverge` function to calculate the age at which the QSD is within 5%.

```{r}
qsdConverge(m1,startLife=SL,conv=0.05)
```
This tells us that in this case, convergence is almost reached at age = `r qsdConverge(m1,startLife=SL,conv=0.05)`.
